---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(magrittr)
dat_raw <- 
  readxl::read_excel("..\\KenaiTrout\\data-raw\\Copy of All fish 2018 Mid River RBT.xlsx",
                     sheet = "All fish 2018 Mid River RBT",
                     col_names = c("date0", "boat", "crew", "method", "time0", "fishnum", "loc", "lat", "long", "lg", "recap", "tag", "ad", "sex0", 
                                   "mat0", "fate0", "comment1_0", "comment2_0", "comment3_0"),
                     col_types = c(rep("text", 5), "numeric", "text", rep("skip", 2), "numeric", "logical", "numeric", "logical", "numeric", 
                                   "text", "numeric", rep("text", 3)),
                     skip = 1) %>%
  dplyr::mutate(date = as.Date(as.character(paste0("0", date0)), "%m%d%Y"), 
                week = as.character(as.character(date, format = "%W")),
                nchar0 = nchar(time0),
                time00 = ifelse(nchar0 == 5, paste0(0, time0), time0),
                #dt = as.POSIXct(as.POSIXlt(time00, format = "%H%M%S"), tz = "America/Anchorage"),
                dt = as.POSIXct(paste0(0, date0, ":", time00), format = "%m%d%Y:%H%M%S", tz = "America/Anchorage"),
                fate = plyr::mapvalues(fate0, from = c(0, 1), to = c("censor", "rel")),
                sex = plyr::mapvalues(sex0, from = c(0, 1, 2), to = c("U", "M", "F")),
                mat = plyr::mapvalues(mat0, from = c(0, 1, 2, 3), to = c(NA, "pre-spawn", "spawning", "post-spawn")),
                comment0 = paste0(ifelse(!is.na(comment1_0), paste0(comment1_0, ifelse(is.na(comment2_0) & is.na(comment3_0), "", ", ")), ""), 
                                 ifelse(!is.na(comment2_0), paste0(comment2_0, ifelse(is.na(comment3_0), "", ", ")), ""),
                                 ifelse(is.na(comment3_0), "", comment3_0)),
                comment = ifelse(comment0 == "", NA, comment0)) %>%
  dplyr::select(-dplyr::ends_with("0")) %>%
  dplyr::arrange(date, boat, fishnum)
```
## 2018 Middle River Rainbow Trout data error check
  
Here are some possible errors. Included most errors in case they affect later inconsistencies. We will likely have to go through this a few times, for example there are no length comparisons at this point and I think we found a few bad tag numbers by looking at large length discrepancies last year. For the time being I stopped after merging the 2017 and 2018 middle river data sets because there are a large number of tags marked as recaps that don't have an initial capture in either data set.
  
1) There are 34 identical rows in the data set. Here is one copy of each duplicated row.  
```{r}
dup_id <- duplicated(dat_raw) 
sum(dup_id) #36 duplicate rows!
dup <- dat_raw[dup_id, ]
dat_raw <- dat_raw[!dup_id, ] # 2495 before, 2459 after
print(dup[, c("date", "boat", "fishnum", "tag", "loc", "lg", "recap", "ad", "fate", "sex", "mat")], n = 50)
```

2) Quite a few times do not make sense. This is a small point but the times are used to pick the first capture when a fish is captured several times within the same event.  
```{r}
hist(as.numeric(format(dat_raw$dt, format = "%H")), main = "hour of catch")
```

3) Two lengths are obviously wrong. Can we fix those? If so there are some others that are lg = 0.  
```{r}
#table(dat_raw$lg[dat_raw$lg <= 150], useNA = "ifany")
dat_raw[dat_raw$lg <= 150 & dat_raw$lg > 0, c("date", "boat", "fishnum", "tag", "loc", "lg", "recap", "ad", "fate", "sex", "mat")]
```

5) One row has an NA for recap.  
```{r}
  dat_raw[dat_raw$tag == 1113, c("date", "boat", "fishnum", "tag", "loc", "lg", "recap", "ad", "fate", "sex", "mat")]
```

5) Twenty one records with 0 for tag number. Most of these look like morts but one fish is not censored and does not have an ominous comment (row # 14).  
```{r}
  dat_raw[dat_raw$tag == 0, c("date", "boat", "fishnum", "tag", "loc", "lg", "recap", "ad", "fate", "comment")] %>% print(n = 100)
  dat_raw <- dat_raw[dat_raw$tag != 0, ] #2459 records before, 2438 after
```

6) Tag 36 has a void note (first record) but looks to be deployed while the last record below says void but I don't see a missing tag number (I thought it referred to tag 1451 but it looks to be deployed a few day later.
```{r}
dat_raw[grepl("VOID", dat_raw$comment), c("date", "boat", "fishnum", "tag", "loc", "lg", "recap", "ad", "fate", "comment")]
```

7) I assume 9999, 99999, and 999999 are codes for "no tag" but note that 2 of those are marked as recaps but are not marked as ad clips.
```{r}
dat_raw[dat_raw$tag %in% c("9999", "99999", "999999"), c("date", "boat", "fishnum", "tag", "loc", "lg", "recap", "ad", "fate", "comment")] # looks like 9999, 99999, and 999999 are dummy tag numbers, delete (7) note 2 are marked recap
dat_raw <- dat_raw[!(dat_raw$tag %in% c("9999", "99999", "999999")), ] #2431 remain
    #dat_raw[dat_raw$tag != 0 & dat_raw$fate == "censor", ] %>% print(n = 100) # save the recaps, delete the rest (12)
      dat_raw <- dat_raw[!(dat_raw$fate == "censor" & dat_raw$recap == 0), ] #2419 remain
      
#Look for fish captured twice in the same week
same_event <-
  dat_raw %>%
  dplyr::group_by(tag, week) %>%
  dplyr::summarise(recap = max(recap), num = n()) %>%
  dplyr::filter(recap == 1 & num > 1)

del_same_event <- 
  dat_raw[dat_raw$tag %in% same_event$tag, ] %>% 
  dplyr::group_by(tag, week) %>%
  dplyr::arrange(tag, week, dt) %>%
  dplyr::filter(dplyr::row_number() != 1) %>%
  dplyr::select(tag, week, dt, boat)
dat_raw <- dplyr::anti_join(dat_raw, del_same_event, by = c("tag", "week", "dt", "boat")) #2419 before, 2377 remain
```

8) 16 tags are in the data twice but neither record is marked as a recap.
```{r}
#Find and correct tags recorded more than once but without a recap flag
dup_norecap <-
  dat_raw %>%
  dplyr::group_by(tag) %>%
  dplyr::summarise(n = n(), recap = sum(recap), lg_range = range(lg)[2] - range(lg)[1]) %>%
  dplyr::filter(recap < (n - 1) & n > 1)
dup_norecap
```

9) There are 177 tag numbers that are marked as a recap but not in the data set twice. You would hope that all of those are 2017 tags. When I check there are 121 I do find in the 2017 data set but 56 that I don't find in the 2017 data set. Here are those tag numbers.
```{r}
#find and correct tags recorded once but marked as a recap
recap_nodup <-
  dat_raw %>%
  dplyr::group_by(tag) %>%
  dplyr::summarise(n = n(), recap = max(recap)) %>%
  dplyr::filter(recap == 1 & n == 1)
#recap_nodup #177 tags marked recap without finding a duplicate
in17dat <- intersect(recap_nodup$tag, KenaiTrout::CH_17$tag)
notin17dat <- setdiff(recap_nodup$tag, KenaiTrout::CH_17$tag)
notin17dat
```

```{r}

```
