---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(magrittr)
rawdat <- readr::read_csv("H:\\My Documents\\Kenai Trout\\Trout2017_MIDDLE RIVER_nov30.csv",
                          col_names = c("date", "rm", "fl", "tag", "sex", "mat", "ad", "tloss", "recap", "hook", "parasite", "fate", "temp", "tl"),
                          col_types = c("ciiciiiiiiiini")) #text", rep("numeric", 2), "text",  rep("numeric", 10)))
rawdat

table(as.Date(rawdat$date, "%Y%m%d"), useNA = "ifany") #Last 2 dates have month and day switched
table(rawdat$rm, useNA = "ifany")
```

Please look at these discrepancies in the CSV data and make any needed corrections.  Some of these are easy enough for me to fix but I thought you should see them so they can be fixed in the main file. Others will change our capture histories so we need to make sure those are all correct. 

1) The month and day are switched on the last 2 sampling days.
```{r}
table(as.Date(rawdat$date, "%Y%m%d"), useNA = "ifany") #Last 2 dates have month and day switched
```

2) Two records show tag loss but both show a tag number and only one is marked as a recap.  I assume these are fish that were captured with adclip and tagging scars but no tag and then retagged on release?
```{r}
rawdat[rawdat$tloss %in% TRUE,] #2 lost tags, one marked as recap one not.
```

3) Probably cleaner to standardize temperature in the csv file.
```{r}
hist(rawdat$temp) #some temps may be in C.

dup_norecap <-
  rawdat[rawdat$tag != 0, ] %>% #9 tag numbers in twice without being marked as recap
  dplyr::group_by(tag) %>% 
  dplyr::summarise(n = n(), recap = max(recap), fl_range = range(fl)[2] - range(fl)[1]) %>% 
  dplyr::filter(recap == 0 & n > 1)
```

4) There are `r dim(dup_norecap)[1]` records where there are 2 entries for a tag number but neither is marked as a recap.  One of those looks like an actual recap to me, the rest probably errors and some may represent errors for several records surrounding the tag number shown below.
```{r}
dup_norecap

recap_nodup <-
  rawdat[rawdat$tag != 0, ] %>% #1 marked as recap with no duplicate record
  dplyr::group_by(tag) %>% 
  dplyr::summarise(n = n(), recap = max(recap)) %>% 
  dplyr::filter(recap == 1 & n == 1)
```

5) There are `r dim(recap_nodup)[1]` records where there is only 1 entry for a tag number but it is marked as a recap.  
```{r}
recap_nodup

dat <- rawdat %>%
        dplyr::mutate(date = as.Date(ifelse(!(gsub("^\\d{4}(\\d{2})\\d{2}", "\\1", date) %in% c("05", "06")),
                                            paste0("201706", gsub("^\\d{4}(\\d{2})\\d{2}", "\\1", date)),
                                            date),
                                     "%Y%m%d"),
                      fl = ifelse(fl == 0, NA, fl),
                      tag = as.numeric(ifelse(tag == "0", NA, tag)),
                      sex = plyr::mapvalues(sex, from = c(0, 1, 2), to = c("U", "M", "F")),
                      mat = as.logical(mat),
                      ad = as.logical(ad),
                      tloss = as.logical(tloss),
                      recap = as.logical(recap),
                      hook = as.logical(hook),
                      parasite = as.logical(parasite),
                      fate = plyr::mapvalues(fate, from = c(0, 1), to = c("rel", "mort")),
                      temp = ifelse(temp < 20, temp * 9 / 5 + 32, temp)
                      )
dat$fate[dat$tag %in% "6044" & dat$date == "2017-05-31"] <- "rel"
dat$recap[dat$tag %in% "6044" & dat$date == "2017-06-06"] <- TRUE

#examine recap data
dat$event <- factor(format(dat$date, "%W"),labels = as.character(1:6))
badmrdat <- is.na(dat$tag) | (dat$fate %in% "mort" & dat$recap == FALSE)
dat_mr <- dat[!badmrdat, ]
recaps <- unique(dat_mr$tag[dat_mr$recap == TRUE])

same_strata <-
  dat_mr %>% 
  tibble::rowid_to_column() %>%
  dplyr::group_by(tag, event) %>% 
  dplyr::summarise(recap = max(recap), num = n(), fl_range = max(fl) - min(fl), row_range = max(rowid) - min(rowid)) %>% 
  dplyr::filter(recap == 1 & num > 1)
```

5) A few records are marked as recaps but were captured during the same time event (week).  These are fine as is but there are a few reasons to be suspicious.  Tag `r same_strata$tag[which(same_strata$fl_range == max(same_strata$fl_range, na.rm = TRUE))]` has a large discrepancy between measured lengths and tag `r same_strata$tag[which(same_strata$row_range == min(same_strata$row_range, na.rm = TRUE))]` has a small number of rows between records.
```{r}
same_strata
```

6) There are also some recaps with very different lengths that should be verified if possible.
```{r}
fill <-
dat_mr[dat_mr$tag %in% recaps, c(1, 3:6, 10:11, 14:15)] %>%
  dplyr::arrange(tag, date) %>%
  dplyr::group_by(tag) %>%
  dplyr::summarise(fl_range = max(fl) - min(fl),
                   fl = mean(fl, na.rm = TRUE),
                   sex = as.character(list(unique(sex[mat == TRUE])[!is.na(unique(sex[mat == TRUE]))])),
                   hook = as.logical(ifelse(is.na(dplyr::last(hook)), max(hook, na.rm = TRUE), dplyr::last(hook))),
                   parasite = as.logical(ifelse(is.na(dplyr::last(parasite)), max(parasite, na.rm = TRUE), dplyr::last(parasite))))
data.frame(fl_range = fill$fl_range) %>% dplyr::group_by(fl_range) %>% dplyr::summarise(num = n())
```
