---
title: "2018 UR estimate"
author: "Adam Reimer"
date: "October 30, 2019"
output: html_document
---

```{r setup, include=FALSE}
library(RMark)
library(magrittr)
library(ggplot2)
# get/prep data
CH_UR18 <- readRDS("..\\data\\CH_UR18.rds")
CH_UR18$lg[is.na(CH_UR18$lg)] <- mean(CH_UR18$lg, na.rm = TRUE)
```

## Assumptions

One thing we might want to change in this report is the assumptions that we list. Instead of the usual 5 assumptions we could go with these 3:    
1. The population is closed to additions (via birth, immigration or growth) and loses (via death and emigration) during the study period,  
2. Marks are neither lost nor overlooked by the investigator, and   
3. Capture probabilities are modeled appropriately.  
  
This list of assumption is probably better (that our standard list about equal probabilities of capture) because we are explicitly modeling capture probability so saying that we assume it is equal for all fish does not make too much sense.

## Data Summaries
The capture history shows substantial variation in the number of fish captured each week, which suggests a model where probability of capture varies by time strata. In the 2009 report you guys tied this variation to discharge, but that relationship is not obvious this year. We do see the percentage of recaptures increasing throughout the study, although not perfectly. This is suggestive of a closed population, but we will explicitly test the closure assumption later.
  
Table 1.- Capture History of Upper Kenai River rainbow trout at least 200mm FL, 2 July--8 August, 2018.
```{r, echo = FALSE}
#Capture History
tab1 <-
  CH_UR18["ch"] %>% 
  tidyr::separate(ch, into = paste0("e", 1:6), sep = 1:5) %>%
  dplyr::mutate_all(.funs = as.numeric) %>%
  dplyr::mutate(n1 = e1,
                n2 = ifelse(e1 == 0, e2, 0),
                n3 = ifelse(e1 + e2 == 0, e3, 0),
                n4 = ifelse(e1 + e2 + e3 == 0, e4, 0),
                n5 = ifelse(e1 + e2 + e3 + e4 == 0, e5, 0),
                n6 = ifelse(e1 + e2 + e3 + e4 + e5 == 0, e6, 0),
                r1 = 0,
                r2 = ifelse(e1 == 1, e2, 0),
                r3 = ifelse(e1 + e2 >= 1, e3, 0),
                r4 = ifelse(e1 + e2 + e3 >= 1, e4, 0),
                r5 = ifelse(e1 + e2 + e3 + e4 >= 1, e5, 0),
                r6 = ifelse(e1 + e2 + e3 + e4 + e5 >= 1, e6, 0),
                a1 = 0,
                a2 = e1,
                a3 = e1 + n2,
                a4 = e1 + n2 + n3,
                a5 = e1 + n2 + n3 + n4,
                a6 = e1 + n2 + n3 + n4 + n5) %>%
  tidyr::gather(name, count) %>%
  dplyr::mutate(stat = gsub("^(.)\\d", "\\1", name),
                event = gsub("^.(\\d)", "\\1", name)) %>%
  dplyr::group_by(stat, event) %>%
  dplyr::summarise(count = sum(count)) %>%
  tidyr::spread(stat, count) %>%
  dplyr::select(capture = e, new = n, recap = r, at_large = a) %>%
  dplyr::mutate(pcap = recap / capture,
                plarge = recap / at_large) %>%
  setNames(c("Captured", "New tags", "Recaptures", "At large", "Recaptures / Captures", "Recaptures / At Large")) %>%
  t()
    
knitr::kable(tab1, col.names = paste0("Event ", 1:6), digits = 3) 
```
  
```{r, echo = FALSE, fig.cap="Figure 1.- Kenai River gauge height at Cooper Landing, 2 July--8 August, 2018."}
read.delim("..\\data-raw\\URlevel_18.txt", 
           header = FALSE, 
           col.names = c("agency", "site", "datetime", "zone", "height", "P"), 
           skip = 28) %>%
  dplyr::mutate(date = as.Date(datetime, "%Y-%m-%d %H:%M")) %>%
  dplyr::filter(date <= as.Date("2018-08-08")) %>%
  dplyr::group_by(date) %>%
  dplyr::summarise(height = mean(height)) %>%
  ggplot(aes(x = date, y = height)) +
    geom_line() +
    annotate("rect", xmin = as.Date("2018-07-02"), xmax = as.Date("2018-07-05"), ymin = -Inf, ymax = Inf, alpha = 0.2) +
  annotate("rect", xmin = as.Date("2018-07-09"), xmax = as.Date("2018-07-11"), ymin = -Inf, ymax = Inf, alpha = 0.2) +
  annotate("rect", xmin = as.Date("2018-07-16"), xmax = as.Date("2018-07-18"), ymin = -Inf, ymax = Inf, alpha = 0.2) +
  annotate("rect", xmin = as.Date("2018-07-23"), xmax = as.Date("2018-07-25"), ymin = -Inf, ymax = Inf, alpha = 0.2) +
  annotate("rect", xmin = as.Date("2018-07-30"), xmax = as.Date("2018-08-01"), ymin = -Inf, ymax = Inf, alpha = 0.2) +
  annotate("rect", xmin = as.Date("2018-08-06"), xmax = as.Date("2018-08-08"), ymin = -Inf, ymax = Inf, alpha = 0.2) +
  labs(y = "Gauge Height", x = "Date") +
  scale_x_date(date_breaks = "week", labels = function(x){format.Date(x, "%B-%d")})
```
  
I first ran this analysis with all six weekly events and found recapture rate differed by river section. In the May 14 version of this summary I included a covariate (location of initial capture) for probability of capture as an incomplete solution. I was never really happy with that however, for one thing fish migrate into different areas and it also complicates the analysis without benefit. Upon revisiting the analysis I realized that probability of capture only differed by location in one week (July 23-25). Because the weekly sample sizes were so large I tried rerunning the analysis without the June 23-25 samples. If we take this step the analysis is streamlined and we still pass precision objectives.
  
Table 2.- Numbers and proportion of rainbow trout at least 200mm FL tagged and recaptured by river section in the upper Kenai River index area, 2 July--8 August, 2018.
```{r, echo = FALSE}
#format data for location based tables
temp <-
  CH_UR18 %>% 
  tidyr::separate(ch, into = paste0("e", 1:6), sep = 1:5) %>%
  tidyr::separate(lh, into = paste0("l", 1:6), sep = 1:5) %>%
  dplyr::mutate_all(.funs = as.numeric) %>%
  dplyr::mutate(n1 = l1,
                n2 = ifelse(e1 == 0, l2, 0),
                n3 = ifelse(e1 + e2 == 0, l3, 0),
                n4 = ifelse(e1 + e2 + e3 == 0, l4, 0),
                n5 = ifelse(e1 + e2 + e3 + e4 == 0, l5, 0),
                n6 = ifelse(e1 + e2 + e3 + e4 + e5 == 0, l6, 0),
                r1 = 0,
                r2 = ifelse(e1 == 1 & e2 == 1, l2, 0),
                r3 = ifelse(e1 + e2 >= 1 & e3 == 1, l3, 0),
                r4 = ifelse(e1 + e2 + e3 >= 1 & e4 == 1, l4, 0),
                r5 = ifelse(e1 + e2 + e3 + e4 >= 1 & e5 == 1, l5, 0),
                r6 = ifelse(e1 + e2 + e3 + e4 + e5 >= 1 & e6 == 1, l6, 0)) %>%
  dplyr::select(tag, lg, dplyr::starts_with("n"), dplyr::starts_with("r")) %>%
  tidyr::gather(event, loc, - tag, -lg) %>%
  dplyr::filter(loc != 0) %>%
  dplyr::arrange(tag) %>%
  dplyr::mutate(class = ifelse(grepl("n", event), "cap", "recap"),
                lg_group = cut(lg, breaks = c(0, 199, 249, 299, 349, 399, 449, 499, 900)))

tab4 <- 
  dplyr::group_by(temp, class, loc) %>%
  dplyr::summarise(n = dplyr::n()) %>%
  tidyr::spread(class, n) %>%
  dplyr::mutate(p1 = recap / (cap + recap),
                p2 = recap / sum(recap),
                p3 = cap / sum(cap))
knitr::kable(tab4, col.names = c("River Section", "New tags", "Recaptures", "Proportion Recaptured", "Proportion of all Recaptures", "Proportion of all Captures"))
chisq.test(tab4[, c("cap", "recap")])

#movement both up and downstream
#population home range is likley larger than the sampling footprint
temp2 <-
  temp %>%
  dplyr::mutate(class2 = ifelse(class == "cap", class, event)) %>%
  dplyr::select(-event, -class) %>%
  dplyr::group_by(tag) %>%
  tidyr::spread(class2, loc)

tab5 <-
  sapply(1:3, function(x){
    dat <- temp2[temp2$cap == x, ]
    table(c(dat$r2, dat$r3, dat$r4, dat$r5, dat$r6))
  }) %>%
  as.data.frame() %>%
  dplyr::mutate(total = V1 + V2 + V3,
                out = total - diag(matrix(c(V1, V2, V3), 3, 3)),
                p = out / total)
```
  
We see a small amount of mixing between river sections. In this case we seem to have both up and downstream movement, at a moderate level. Of all the fish recaptured `r format(sum(tab5$out) / sum(tab5$total), digits = 3)` were recaptured in a different section than they were marked. It seems likely that fish near the boundaries of our study area migrate across it and the home range of our sampled fish is likely larger than the geographic extent of the index area.
  
Table 3.- Movement of rainbow trout at least 200mm FL between river sections in the Upper Kenai River index area, 2 July--8 August, 2018. 
```{r, echo = FALSE}
rownames(tab5) <- c("Captured in 1", "Captured in 2", "Captured in 3")
knitr::kable(tab5, col.names = c("1", "2", "3", "Total recaptures", "# out", "Proportion out"))
```
  
Given the results of the 2009 study we suspect probability of capture may differ by length. The size distribution of fish captured differs by event, which means length driven differences in probability of capture could affect our abundance estimate.

```{r, echo = FALSE, fig.cap = "Figure 2.- Cumulative length distributions of captured rainbow trout during each event in the upper Kenai River index area, 2 July--8 August, 2018."}
CH_UR18$lg[is.na(CH_UR18$lg)] <- mean(CH_UR18$lg, na.rm = TRUE)
lg <- 
  CH_UR18 %>% 
  tidyr::separate(ch, into = paste0("e", 1:6), sep = 1:5) %>%
  dplyr::select(dplyr::starts_with("e"), lg) %>%
  dplyr::mutate_at(.vars = dplyr::vars(dplyr::starts_with("e")), .funs = list(~ as.numeric(.) * lg)) %>%
  dplyr::select(-lg) %>%
  tidyr::gather(event, lg) %>%
  dplyr::filter(lg != 0)
ggplot(lg, aes(x = lg, color = event)) +
  stat_ecdf() +
  ylab("Cumulative Proportion") +
  scale_x_continuous(name = "Fork Length (mm)", breaks = seq(200, 600, 50))
kSamples::ad.test(lapply(paste0("e", 1:5), function(x){lg$lg[lg$event == x]}))
```
  
Recapture rates also increases for the larger length groups, although the pattern is weaker than observed in 2009. As an aside the notation (199,299] should be read as 200-299 because the rounded bracket indicates the range does not include the adjacent number while the square bracket indicates the range does include the adjacent number.
  
Table 4.- Number of rainbow trout captured, recaptured and the proportion recaptured in the Upper Kenai River index area by 100mm length groups, 2 July--8 August, 2018. 
```{r, echo = FALSE}
tab6 <- 
  dplyr::group_by(temp, class, lg_group) %>%
  dplyr::summarise(n = n()) %>%
  tidyr::spread(class, n) %>%
  dplyr::mutate(total = (cap + recap),
                p1 = recap / total)
knitr::kable(tab6, col.names = c("Length group (mm)", "New tags", "Recaptures", "Total", "Proportion Recaptured"))
chisq.test(tab6[, c("cap", "recap")])
```
  
To directly asses the closure assumption I used the POPAN modeling framework we have used on middle river rainbows. Probability of capture was assumed to vary with FL group and time. For survival I considered two structures; one with an estimated constant survival and a second where survival is assumed 100%. The second structure is consistent with the closure assumption. For entrance probability I considered four structures; one where probability of entrance varies with time, a second where probability of entrance is constant, a third where fish only entered during the last event, and one where all of the fish were assumed to be within the study area prior to the first event. The last assumption is consistent with the closure assumption. In Mark we are able to run all combinations of these parameter structures and decide which combinations are best supported by the data. Four models accounted for 95%+ of AIC weight. Three where consistent with the closure assumption. The forth model include the possibility that the closure assumption violated by migration during the last event but those models are not supported by the data. I see that the 2009 report had a lot of upstream migrating fish, but we don't see that this time around. 
  
## Candidate models
In an effort to simplify model selection I'll present the candidate models in 2 stages. The first stage contains basic closed multi-event mark recapture models; M0 (constant probability of capture), Mt (time varying probability of capture), Mh (individual heterogeneity in probability of capture), and Mb (behavioral effects in probability of capture). Individual heterogeneity was modeled in three ways; a general version that treats probability of capture as a mixture of two distributions, a version that used FL as a individual covariate, and a version that uses the river section of initial capture as a individual covariate. The results strongly support time varying capture probability.
```{r}
closed_results1 <- readRDS("..\\scripts\\2018upper\\closed_results1.rds")
knitr::kable(closed_results1$model.table,
             digits = 3)
```
  
In the second stage of model selection I combined some of the lesser preforming models with Mt. The top performing model allowed probability of capture to vary by time and fish size. The second model is similar but adds a behavior effect. Because parameter estimates for the behavior effect include zero, this model should not be considered.
```{r}
closed_results2 <- readRDS("..\\scripts\\2018upper\\closed_results2.rds")
bestmod2 <- as.numeric(rownames(closed_results2$model.table)[1:2])
knitr::kable(closed_results2$model.table[, -which(names(closed_results2$model.table) %in% c("p", "c"))],
             digits = 3)
```
  
Table 5.- Parameter estimates for the top preforming abundance estimation model, rainbow trout at least 200mm FL between river sections in the Upper Kenai River index area, 2 July--8 August, 2018.
```{r, echo = FALSE}
knitr::kable(closed_results2[[bestmod2[1]]]$results$beta, digits = 3)
```
  
```{r, echo = FALSE, fig.height=6, fig.width=8, fig.cap = "Figure 3.- Predicted probability of rainbow trout capture vs. fork length for each event and river section of initial capture in the Upper Kenai River index area, 2 July--8 August, 2018."}
plot_dat <- expand.grid(lg = seq(200,600,25), index = 1:6)
event_labs <- c("July 2-5", "July 9-11", "July 16-18", "July 23-25", "July 30-Aug. 1", "Aug 6-8")
covariate.predictions(closed_results2[[bestmod2[1]]], data = plot_dat, indices = c(1, 6))$estimates %>%
  dplyr::mutate(event = factor(index, labels = event_labs)) %>%
  ggplot(aes(x = lg, ymin = lcl, ymax = ucl)) +
    geom_ribbon(alpha = 0.25, linetype = 0) +
    geom_line(aes(y = estimate)) +
    facet_grid(.~event) +
    labs(y = "Probability of Capture", x = "Total Length")
``` 
  
Table 6.- Real parameter estimates for the top preforming abundance estimation model, rainbow trout at least 200mm FL between river sections in the Upper Kenai River index area, 2 July--8 August, 2018.
```{r, echo = FALSE}
knitr::kable(closed_results2[[bestmod2[1]]]$results$real, digits = 3)
```
  
Table 7.- Estimated abundance of rainbow trout at least 200mm FL between river sections in the Upper Kenai River index area, 2 July--8 August, 2018.
```{r, echo = FALSE}
knitr::kable(closed_results2[[bestmod2[1]]]$results$derived, digits = 3)
```
  
Table 8.- Estimated abundance and proportion of rainbow trout by fork length group in the Upper Kenai River index area, 2 July--8 August, 2018.
```{r, echo = FALSE}
#Weighted length comp
alpha <- closed_results2[[bestmod2[1]]]$results$beta$estimate[grepl("p:\\(|p:time\\d$", rownames(closed_results2[[bestmod2[1]]]$results$beta))]
beta <- closed_results2[[bestmod2[1]]]$results$beta$estimate[grepl("p:lg|p:time\\d:lg", rownames(closed_results2[[bestmod2[1]]]$results$beta))] 
num <- 
  CH_UR18 %>% 
  dplyr::mutate(event = regexpr("1", ch),
                c_ik = ifelse(event == 1,
                              exp(alpha[1] + beta[1] * lg)/(1 + exp(alpha[1] + beta[1] * lg)),
                              exp(alpha[1] + alpha[event] + beta[1] * lg + beta[event] * lg) /
                                 (1+ exp(alpha[1] + alpha[event] + beta[1] * lg + beta[event]))),
                lg_bin = cut(lg, breaks = seq(200, 600, 50), right = FALSE)) %>%
  dplyr::group_by(event, lg_bin) %>%
  dplyr::summarize(num_p = sum(1 / c_ik), n_ij = n())
dem <- 
  num %>% 
  dplyr::summarise(dem_p = sum(num_p),
                   n_i = sum(n_ij))
temp_age <- 
  dplyr::left_join(num, dem, by = "event") %>%
  dplyr::ungroup() %>%
  dplyr::mutate(p_ij = num_p/dem_p,
                w_i = n_i / sum(dem$n_i),
                lg_bin2 = ifelse(lg_bin %in% c("[200,250)", "[250,300)"), "[200,300)", "[300,600)"))

temp_age %>%
  dplyr::group_by(lg_bin) %>%
  dplyr::summarise(n_j = sum(n_ij),
                   pj_raw = n_j / sum(dem$n_i),
                   se_pj_raw = sqrt(pj_raw * (1 - pj_raw) / (sum(dem$n_i) - 1)),
                   p_j = sum(w_i * p_ij),
                   se_pj = se_pj_raw,
                   N_j = p_j * closed_results2[[bestmod2[1]]]$results$derived$'N Population Size'$estimate,
                   se_N_j = sqrt(closed_results2[[bestmod2[1]]]$results$derived$'N Population Size'$estimate^2 * se_pj^2 +
                                   p_j^2 * closed_results2[[bestmod2[1]]]$results$derived$'N Population Size'$se^2 -
                                   se_pj^2 * closed_results2[[bestmod2[1]]]$results$derived$'N Population Size'$se^2)) %>%
  knitr::kable(digits = 3)
```

```{r}
#again for >300mm
temp_age %>%
  dplyr::group_by(lg_bin2) %>%
  dplyr::summarise(n_j = sum(n_ij),
                   pj_raw = n_j / sum(dem$n_i),
                   se_pj_raw = sqrt(pj_raw * (1 - pj_raw) / (sum(dem$n_i) - 1)),
                   p_j = sum(w_i * p_ij),
                   se_pj = se_pj_raw,
                   N_j = p_j * closed_results2[[bestmod2[1]]]$results$derived$'N Population Size'$estimate,
                   se_N_j = sqrt(closed_results2[[bestmod2[1]]]$results$derived$'N Population Size'$estimate^2 * se_pj^2 +
                                   p_j^2 * closed_results2[[bestmod2[1]]]$results$derived$'N Population Size'$se^2 -
                                   se_pj^2 * closed_results2[[bestmod2[1]]]$results$derived$'N Population Size'$se^2)) %>%
  knitr::kable(digits = 3)
```
